MSB1015_Assignment 2

### 2019-09-28 MSB1015_Assignment2_SuzannetenHage.Rmd
Update MSB1015_Assignment2_SuzannetenHage.Rmd
updated descriptions
added null model  
added PLS
added visualizations of data and PLS results

**0. Load packages**
If you haven't installed the packages on your computer yet, please uncomment the 
install.packages lines.

```{r}
# install.packages("WikidataQueryServiceR")
library("WikidataQueryServiceR")
#install.packages('rJava')
library('rJava')
#install.packages('rcdk')
library('rcdk')
#install.packages('pls')
library('pls')

@@

**1. Get the data from WikiData.** 
In this section I use the WikidataQueryServiceR package to obtain all simple alkines with known boiling points from WikiData.

```{r}
data = query_wikidata('SELECT ?comp ?compLabel ?bp ?bpUnit ?bpUnitLabel ?smiles WHERE {   
    ?comp wdt:P31/wdt:P279* wd:Q41581 ;
    p:P2102 [ps:P2102 ?bp ;           
            psv:P2102/wikibase:quantityUnit  
            ?bpUnit         
            ] ;
    wdt:P233 ?smiles.
    SERVICE wikibase:label { bd:serviceParam wikibase:language "[AUTO_LANGUAGE],en". } 
    } '
)

@@

**3.Quantify the alkane structures**

@@

#Get descriptors
parsed_smiles <- parse.smiles(data$smiles, kekulise=TRUE)

descNames <- c(
'org.openscience.cdk.qsar.descriptors.molecular.APolDescriptor',
'org.openscience.cdk.qsar.descriptors.molecular.WienerNumbersDescriptor'
)

descriptors <- eval.desc(parsed_smiles, descNames)
```

**4. Data pre-processing**
Not all boiling points in the database are in kelvin, which is the SI unit for temperature. Before I continue, I make sure all temperatures are in kelvin.

```{r}
#Find all rownumbers of the values in Celcius. 
celcius <-  which(grepl('degree Celsius', data$bpUnitLabel))
for (i in 1:length(celcius)) {
  #Convert the value to kelvin
  data$bp[celcius[i]] <- data$bp[celcius[i]] + 273.15
  #Convert the label to kelvin. 
  data$bpUnitLabel[celcius[i]] <- 'kelvin'
}

@@

>>>>

**0. Load packages**
If you haven't installed the packages on your computer yet, please uncomment the install.packages lines before running the code below.

# install.packages("WikidataQueryServiceR")
library("WikidataQueryServiceR")
#install.packages('rJava')
library('rJava')
#install.packages('rcdk')
library('rcdk')
#install.packages('caTools')
library('caTools')
#install.packages('pls')
library('pls')
#install.packages('Metrics')
library('Metrics')

@@

**1. Get the data from WikiData.** 
In this section I use the WikidataQueryServiceR package to obtain all simple alkines with known boiling points from WikiData. I get their name, boiling point, boiling point unit and smiles.

data = query_wikidata('SELECT ?comp ?compLabel ?bp ?bpUnit ?bpUnitLabel ?smiles WHERE {   
    ?comp wdt:P31/wdt:P279* wd:Q41581 ;
    p:P2102 [ps:P2102 ?bp ;           
            psv:P2102/wikibase:quantityUnit  
            ?bpUnit         
            ] ;
    wdt:P233 ?smiles.
    SERVICE wikibase:label { bd:serviceParam wikibase:language "[AUTO_LANGUAGE],en". } 
    } '
)

@@

**2. Get the descriptors of the alkane structures**
I get descriptors for all alanes from CDK based on their smiles. These descriptors will be used later as predictors in the PLS model.

@@

#Reformatting of smiles into correct format for CDK. Kekulise checks for electrons, and prevents parsing of incorrect smiles. 
parsed_smiles <- parse.smiles(data$smiles, kekulise=TRUE)

#Determine which descriptors to get. 
descNames <- c(
'org.openscience.cdk.qsar.descriptors.molecular.APolDescriptor',
'org.openscience.cdk.qsar.descriptors.molecular.WienerNumbersDescriptor')

#Get descriptors
descriptors <- eval.desc(parsed_smiles, descNames)

**3. Data pre-processing**
3a) Not all boiling points in the database are in kelvin, which is the SI unit for temperature. Before I continue, I make sure all temperatures are in kelvin.

#Find all rownumbers of the values in Celcius. 
celcius <-  which(grepl('degree Celsius', data$bpUnitLabel))
for (i in 1:length(celcius)) {
  #Convert the value to kelvin
  data$bp[celcius[i]] <- data$bp[celcius[i]] + 273.15
  #Convert the label to kelvin. 
  data$bpUnitLabel[celcius[i]] <- 'kelvin'
}

@@

3b) To be able to test how well the model I train performs, it is necessary to keep a part of our data aside as testing data. In the following code you can easily change the size of the test data by adjusting percentage_test.

#Percentage of data to go into the test-set. 
percentage_test <- 0.25

#Splitting data into a training and test set. 
set.seed(50) 
sample = sample.split(full_data, SplitRatio = percentage_test)
train = subset(full_data, sample == FALSE)
test  = subset(full_data, sample == TRUE)

**4. Null model**
The null model is the most simple model possible: the mean of the data. This model can be used as a reference to see whether the PLS model performs better.

#Create null_model
null_model <- mean(train$data.bp)

#Get error of null model
error_null_model <- rmse(test$data.bp, null_model)

**5. Partial Least Squares**
Here, I train the PLS model with the training data, and use the model to predict the boiling points of the test data.

#Training of the PLS model. 
PLS_model <- plsr(data.bp ~ (apol + WPATH + WPOL), 
                  data = train, 
                  validation = "CV"
                  )

#Predict the boiling points of the test set. 
predicted.bp <- predict(PLS_model, test)

#Calculate the error. 
error <- RMSEP(PLS_model, newdata = test)

**6. Visualization**

#Plot the predicted data vs. the original data. 
plot(test$data.bp, predicted.bp[,,3], xlim=c(250,1000), ylim=c(250,1000))
